GPU Efficient Texture Based Bezier Curve Evaluation

### 2015-06-25 Paper/Figures/FigTexture1D.tex
Paper: added accuracy diagrams, combined glsl listings

\documentclass{article}
\usepackage{tikz}
\definecolor{darkgreen}{RGB}{0,192,0}
\title{Cartesian closed categories and the price of eggs}
\author{Jane Doe}
\date{September 1994}
\begin{document}

	\usetikzlibrary{arrows.meta}
	\tikzset{>={Latex[width=3mm,length=3mm]}}

	\begin{figure}
		\begin{tikzpicture}[x=12.5cm,y=4.25cm]
			% border
			\draw (0cm,0cm) -- (12.5cm,0cm) -- (12.5cm,4.25cm) -- (0cm,4.25cm) -- (0cm, 0cm);		
			% raw texture, bilinear sampled texture, curve results
		    \node[anchor=south west,inner sep=0] at (0.125cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Texture.png}};
		    \node[anchor=south west,inner sep=0] at (4.25cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Bilinear.png}};
		    \node[anchor=south west,inner sep=0] at (8.375cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Curves.png}};
		    % representative line for where the quadratic bezier curve lives
		    \draw[yellow,->,line width=0.5mm]    (5.25cm,1.125cm) -- (7.25cm,3.125cm);    
		    \draw (0.08,0.25) node[] {\Huge{A}};
		    \draw (0.25,0.25) node[] {\Huge{B}};
		    \draw (0.08,0.75) node[] {\Huge{B}};
		    \draw (0.25,0.75) node[] {\Huge{C}};    		    
		\end{tikzpicture}
		\caption{Left: a 2x2 texture storing control points for a quadratic Bezier curve in each color channel.  Center: The same texture as viewed when using bilinear texture sampling.  The yellow line indicates where texture samples are taken from to evaluate the quadratic Bezier curve. Right: The curves resulting from sampling along the yellow line (Alpha curve ommited).}		
	\end{figure}

>>>>

\documentclass{article}
\usepackage{tikz}
\usepackage{tkz-euclide}
\definecolor{darkgreen}{RGB}{0,192,0}
\title{Cartesian closed categories and the price of eggs}
\author{Jane Doe}
\date{September 1994}
\begin{document}

	\begin{figure}
		\begin{tikzpicture}[x=8cm,y=4cm]
			\tkzInit[xmax=2,ymax=1,xmin=0,ymin=0]
			% border
			\draw (-0.1,-0.2) rectangle (2,1.2);

			\foreach \x in {0,...,4}
				\draw (\x * 0.1 - 0.05,0.45) rectangle (\x * 0.1 + 0.05, 0.55);

		\end{tikzpicture}
		\caption{TODO: this}		
	\end{figure}

### 2015-06-25 Paper/paper.tex
Paper: added accuracy diagrams, combined glsl listings

The De Casteljeau algorithm shows us that we can find points on a Bezier curve by evaluating a hierarchy of linear interpolations.  For example, a quadratic curve is the linear interpolation between the linear interpolation of {$\overline{AB}$} and {$\overline{BC}$}, using the same time $t$ for each interpolation (\autoref{lst:GLSLDeCasteljeau}).

\begin{lstlisting}[caption={GLSL De Casteljeau Algorithm For a Quadratic Bezier Curve}, label={lst:GLSLDeCasteljeau}]
float QuadraticBezier (
  in float t,
  in float A,
  in float B,
  in float C
) {
    return mix(mix(A,B,t), mix(B,C,t), t);
}
\end{lstlisting}

Bilinear interpolation is used when linear sampling is enabled and a texture sample is taking from a 2d texture.  Bilinear interpolation is achieved by linearly interpolating across one axis (say, the X axis) and then interpolating the two results by the other axis (the Y axis).  In bilinear interpolation, you have a $t$ and a $u$ interpolation value, and you have four values to interpolate between $A$,$B$,$C$,$D$ (\autoref{lst:GLSLBilinear}).

\begin{lstlisting}[caption={GLSL implementation of bilinear interpolation between four floats}, label={lst:GLSLBilinear}]
float BilinearInterpolate (
  in float u,
  in float v,
  in float A,
  in float B,
  in float C,
  in float D
) {
    return mix(mix(A,B,u), mix(C,D,u), v);
}
\end{lstlisting}

Comparing the source code for these two operations, we begin to get an idea of how to set up the points in our texture and the u and v coordinates to use such that when performing bilinear sampling, that we will get, as output, points on a quadratic bezier curve (\autoref{fig:decasteljeauBilinear}).

  \begin{figure}
    % bezier curve diagram
    \begin{tikzpicture}[x=4cm,y=4cm]
      % border
      \draw (-0.2,-0.2) -- (1.2,-0.2) -- (1.2,1.2) -- (-0.2,1.2) -- (-0.2,-0.2);
      % 1d quadratic bezier curve with control points 0,1,0
      \draw[scale=1,domain=0:1,smooth,variable=\x,blue,line width=0.5mm] plot ({\x},{2*\x-2*\x*\x});    
      % control polygon / labels
      \draw[red,->]       (0.0,0.0) -- (0.5,1.0);
      \draw[darkgreen,->] (0.5,1.0) -- (1.0,0.0);
      \draw[red]          (0.25,0.5) node[anchor=east] {$\overline{AB}$};
      \draw[darkgreen]    (0.75,0.5) node[anchor=west] {$\overline{BC}$};
      % control point labels
      \draw (0.0,0.0) node[anchor=east]  {\bf{A}};
      \draw (0.5,1.0) node[anchor=south] {\bf{B}};
      \draw (1.0,0.0) node[anchor=west]  {\bf{C}};
      % draw a line from AB->AC for time = 0.75
      \draw[orange,->] (0.3,0.6) -- (0.8,0.4);
      \draw[orange] (0.40,0.55) node[anchor=south west] {$\stackrel{\frown}{ABC}$};
      % draw the point at t=0.6 and label underneath the curve
      \draw[black,fill=black] (0.6,0.48) circle (1mm);
      \draw[black] (0.5,0.45) node[anchor=north] {\bf{t}=0.6};
    \end{tikzpicture}
    % bilinear diagram
    \begin{tikzpicture}[x=4cm,y=4cm]
        % border
        \draw (-0.2,-0.2) -- (1.2,-0.2) -- (1.2,1.2) -- (-0.2,1.2) -- (-0.2,-0.2);
        % box
        \draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- (0,0);
        % representative line for where the quadratic bezier curve lives
        \draw[blue,line width=0.5mm] (0,0) -- (1,1);    
        % corner labels
        \draw (0,0) node[anchor=north east] {\bf{A}};
        \draw (1,0) node[anchor=north west] {\bf{B}};     
        \draw (0,1) node[anchor=south east] {\bf{B}};
        \draw (1,1) node[anchor=south west] {\bf{C}};     
        % x axis interpolation
        \draw[red,->]       (0.0,0.1) -- (1.0,0.1);
        \draw[red]          (0.4,0.1) node[anchor=south] {$\overline{AB}$};
        \draw[darkgreen,->] (0.0,0.9) -- (1.0,0.9);
        \draw[darkgreen]    (0.4,0.9) node[anchor=north] {$\overline{BC}$};
        % y axis interpolation
        \draw[orange,->]    (0.6,0.0) -- (0.6,1.0);
        \draw[orange]     (0.6,0.5) node[anchor=west] {$\stackrel{\frown}{ABC}$};
        % draw the point at t=0.6 and label for it
        \draw[black,fill=black] (0.6,0.6) circle (1mm);
        \draw[black] (0.55,0.6) node[anchor=east] {\bf{t}=0.6};
    \end{tikzpicture} 
    \caption{Left: A quadratic Bezier curve evaluated at {\bf{t}}=0.6 for control points A,B,C using the De Casteljeau algorithm.  Right: The De Casteljeau algorithm using bilinear interpolation.  Note that in this case, the X axis is evaluated before the Y, but the Y axis could be evaluated before the X for the same results.}
    \label{fig:decasteljeauBilinear}
  \end{figure}

@@

  \begin{figure}
    \begin{tikzpicture}[x=12.5cm,y=4.25cm]
      % border
      \draw (0cm,0cm) -- (12.5cm,0cm) -- (12.5cm,4.25cm) -- (0cm,4.25cm) -- (0cm, 0cm);   
      % raw texture, bilinear sampled texture, curve results
        \node[anchor=south west,inner sep=0] at (0.125cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Texture.png}};
        \node[anchor=south west,inner sep=0] at (4.25cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Bilinear.png}};
        \node[anchor=south west,inner sep=0] at (8.375cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Curves.png}};
        % representative line for where the quadratic bezier curve lives
        \draw[yellow,->,line width=0.5mm]    (5.25cm,1.125cm) -- (7.25cm,3.125cm);    
        \draw (0.08,0.25) node[] {\Huge{A}};
        \draw (0.25,0.25) node[] {\Huge{B}};
        \draw (0.08,0.75) node[] {\Huge{B}};
        \draw (0.25,0.75) node[] {\Huge{C}};            
    \end{tikzpicture}
    \caption{Left: a 2x2 texture storing control points for a quadratic Bezier curve in each color channel.  Center: The same texture as viewed when using bilinear texture sampling.  The yellow line indicates where texture samples are taken from to evaluate the quadratic Bezier curve. Right: The curves resulting from sampling along the yellow line (Alpha curve ommited).}   
    \label{fig:texbilcurve}
  \end{figure}

@@

The technique presented in this paper also has numerical differences in practice.  While it's been shown that the multidimensional linear interpolation capabilities can be mathematically equivelant to the Bernstein form as well as the De Casteljeau algorithm, there are mathematical precision issues due to current GPU architectures.

@@

TODO: give the details of the architecture and cite the two papers cited from the other place!

TODO: Show a zoomed in version of curves with this technique, vs the real curve

%-------------------------------------------------------------------------
\section{Implications of Texture Dimensionality}
\label{sec:implicationsoftexturedimensionality}

While this technique is able to be used with any texture dimensionality, there are different characteristics when using textures of different dimensionalities.  Those characteristics will be explored here.

@@

That means that if you want a Bezier curve higher than order 1, you are going to have to take more texture samples and combine the results in the shader program.  While this is more expensive computationally, it does have the benefit of doing more of the work in the shader program, which has higher precision, so the end result will be a higher quality curve.

>>>>

The De Casteljeau algorithm shows us that we can find points on a Bezier curve by evaluating a hierarchy of linear interpolations.  For example, a quadratic curve is the linear interpolation between the linear interpolation of {$\overline{AB}$} and {$\overline{BC}$}, using the same time $t$ for each interpolation.

Bilinear interpolation is used when linear sampling is enabled and a texture sample is taking from a 2d texture.  Bilinear interpolation is achieved by linearly interpolating across one axis (say, the X axis) and then interpolating the two results by the other axis (the Y axis).  In bilinear interpolation, you have a $t$ and a $u$ interpolation value, and you have four values to interpolate between $A$,$B$,$C$,$D$.

\begin{lstlisting}[caption={GLSL implementation of bilinear interpolation and the De Casteljeau algorithm for a quadratic Bezier curve.}, label={lst:GLSLCurves}]
float BilinearInterpolate (
  in float u, in float v,
  in float A, in float B, in float C, in float D
) {
    return mix(mix(A,B,u), mix(C,D,u), v);
}

float QuadraticBezier (
  in float t,
  in float A, in float B, in float C
) {
    return mix(mix(A,B,t), mix(B,C,t), t);
}
\end{lstlisting}

Comparing the source code for these two operations  (\autoref{lst:GLSLCurves}), we begin to get an idea of how to set up the points in our texture and the u and v coordinates to use such that when performing bilinear sampling, that we will get, as output, points on a quadratic bezier curve (\autoref{fig:decasteljeauBilinear}).

  \begin{figure}
    % bezier curve diagram
    \begin{tikzpicture}[x=4cm,y=4cm]
      % border
      \draw (-0.2,-0.2) -- (1.2,-0.2) -- (1.2,1.2) -- (-0.2,1.2) -- (-0.2,-0.2);
      % 1d quadratic bezier curve with control points 0,1,0
      \draw[scale=1,domain=0:1,smooth,variable=\x,blue,line width=0.5mm] plot ({\x},{2*\x-2*\x*\x});    
      % control polygon / labels
      \draw[red,->]       (0.0,0.0) -- (0.5,1.0);
      \draw[darkgreen,->] (0.5,1.0) -- (1.0,0.0);
      \draw[red]          (0.25,0.5) node[anchor=east] {$\overline{AB}$};
      \draw[darkgreen]    (0.75,0.5) node[anchor=west] {$\overline{BC}$};
      % control point labels
      \draw (0.0,0.0) node[anchor=east]  {\bf{A}};
      \draw (0.5,1.0) node[anchor=south] {\bf{B}};
      \draw (1.0,0.0) node[anchor=west]  {\bf{C}};
      % draw a line from AB->AC for time = 0.75
      \draw[orange,->] (0.3,0.6) -- (0.8,0.4);
      \draw[orange] (0.40,0.55) node[anchor=south west] {$\stackrel{\frown}{ABC}$};
      % draw the point at t=0.6 and label underneath the curve
      \draw[black,fill=black] (0.6,0.48) circle (1mm);
      \draw[black] (0.5,0.45) node[anchor=north] {\bf{t}=0.6};
    \end{tikzpicture}
    % bilinear diagram
    \begin{tikzpicture}[x=4cm,y=4cm]
        % border
        \draw (-0.2,-0.2) -- (1.2,-0.2) -- (1.2,1.2) -- (-0.2,1.2) -- (-0.2,-0.2);
        % box
        \draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- (0,0);
        % representative line for where the quadratic bezier curve lives
        \draw[blue,line width=0.5mm] (0,0) -- (1,1);    
        % corner labels
        \draw (0,0) node[anchor=north east] {\bf{A}};
        \draw (1,0) node[anchor=north west] {\bf{B}};     
        \draw (0,1) node[anchor=south east] {\bf{B}};
        \draw (1,1) node[anchor=south west] {\bf{C}};     
        % x axis interpolation
        \draw[red,->]       (0.0,0.1) -- (1.0,0.1);
        \draw[red]          (0.4,0.1) node[anchor=south] {$\overline{AB}$};
        \draw[darkgreen,->] (0.0,0.9) -- (1.0,0.9);
        \draw[darkgreen]    (0.4,0.9) node[anchor=north] {$\overline{BC}$};
        % y axis interpolation
        \draw[orange,->]    (0.6,0.0) -- (0.6,1.0);
        \draw[orange]     (0.6,0.5) node[anchor=west] {$\stackrel{\frown}{ABC}$};
        % draw the point at t=0.6 and label for it
        \draw[black,fill=black] (0.6,0.6) circle (1mm);
        \draw[black] (0.55,0.6) node[anchor=east] {\bf{t}=0.6};
    \end{tikzpicture} 
    \caption{\textit{Left:} A quadratic Bezier curve evaluated at {\bf{t}}=0.6 for control points A,B,C using the De Casteljeau algorithm.  \textit{Right:} The De Casteljeau algorithm using bilinear interpolation.  Note that in this case, the X axis is evaluated before the Y, but the Y axis could be evaluated before the X for the same results.}
    \label{fig:decasteljeauBilinear}
  \end{figure}

@@

  \begin{figure}
    \begin{tikzpicture}[x=12.5cm,y=4.25cm]
      % border
      \draw (0cm,0cm) -- (12.5cm,0cm) -- (12.5cm,4.25cm) -- (0cm,4.25cm) -- (0cm, 0cm);   
      % raw texture, bilinear sampled texture, curve results
        \node[anchor=south west,inner sep=0] at (0.125cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Texture.png}};
        \node[anchor=south west,inner sep=0] at (4.25cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Bilinear.png}};
        \node[anchor=south west,inner sep=0] at (8.375cm,0.125cm) {\includegraphics[width=4cm,height=4cm]{Figure4Curves.png}};
        % representative line for where the quadratic bezier curve lives
        \draw[yellow,->,line width=0.5mm]    (5.25cm,1.125cm) -- (7.25cm,3.125cm);    
        \draw (0.08,0.25) node[] {\Huge{A}};
        \draw (0.25,0.25) node[] {\Huge{B}};
        \draw (0.08,0.75) node[] {\Huge{B}};
        \draw (0.25,0.75) node[] {\Huge{C}};            
    \end{tikzpicture}
    \caption{\textit{Left:} a 2x2 texture storing control points for a quadratic Bezier curve in each color channel.  \textit{Middle:} The same texture as viewed when using bilinear texture sampling.  The yellow line indicates where texture samples are taken from to evaluate the quadratic Bezier curve. \textit{Right:} The curves resulting from sampling along the yellow line (Alpha curve ommited).}   
    \label{fig:texbilcurve}
  \end{figure}

@@

The technique presented in this paper also has numerical differences in practice.  While it's been shown that the multidimensional linear interpolation capabilities can be mathematically equivelant to the Bernstein form as well as the De Casteljeau algorithm, there are mathematical precision issues due to current GPU architectures (\autoref{fig:quickaccuracy}).

@@

  \begin{figure}
    \begin{tikzpicture}[x=12cm,y=4cm]
      % border
      \draw (0cm,0cm) rectangle (12cm,4cm);
      % raw texture, bilinear sampled texture, curve results
        \node[anchor=south west,inner sep=0] at (0cm,0cm) {\includegraphics[width=4cm,height=4cm]{HighQuality.png}};
        \node[anchor=south west,inner sep=0] at (4cm,0cm) {\includegraphics[width=4cm,height=4cm]{HighQualityZoom.png}};
        \node[anchor=south west,inner sep=0] at (8cm,0cm) {\includegraphics[width=4cm,height=4cm]{LowQualityZoom.png}};
        % representative line for where the quadratic bezier curve lives
        \draw[yellow] (0.11,0.35) rectangle (0.21,0.65);

    \end{tikzpicture}
    \caption{\textit{Left:} Sampled curves.  \textit{Middle:} Zooming in on a curve calculated with shader program constants.  \textit{Right:} Zooming in on a curve calculated using the method in this paper}   
    \label{fig:quickaccuracy}
  \end{figure}

TODO: give the details of the architecture imprecision sources and cite the two papers cited from the other place!

%-------------------------------------------------------------------------
\section{Texture Dimensionality}
\label{sec:texturedimensionality}

While this technique is able to be used with any texture dimensionality, there are different characteristics for each choice.  Those characteristics will be explored here.

@@

If you want a Bezier curve higher than order 1, you are going to have to take more texture samples and combine the results in the shader program.  While this is more expensive computationally, it does have the benefit of doing more of the work in the shader program, which has higher precision through use of full float math, so the end result will be a higher quality curve.
